---
title: "SIR ODE using griddleR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{SIR ODE using griddleR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(griddleR)
library(dplyr)
library(ggplot2)
```

Let's set up a simple simulation, that takes in a list of parameters, runs an
SIR ODE, and outputs the sizes of the compartments at different time points:

```{r model}
# given time, compartment sizes, and paramters, compute rates of compartment
# size change
sir_ode <- function(t, state, params) {
  names(state) <- c("S", "I", "R")
  out <- with(c(state, params), {
    N <- sum(state)
    gamma <- 1.0 / infectious_period
    beta <- R0 * infectious_period

    dS <- -beta * S / N * I
    dI <- -dS - I * gamma
    dR <- I * gamma

    c(dS, dI, dR)
  })

  stopifnot(abs(sum(out)) < 1e-6)

  list(out)
}

sir <- function(params) {
  # initialize simulation state
  state0 <- c(S = 1.0 - params$I0, I = params$I0, R = 0.0)
  # select times to output compartment sizes
  times <- seq(0, params$max_time, length.out = params$n_points)
  # run the ODE
  result <- deSolve::ode(state0, times, sir_ode, params)
  # the result is a weird object: it's a matrix, with columns of type deSolve.
  # instead, make this a tibble with simple numeric columns.
  result |>
    as_tibble() |>
    mutate(across(everything(), as.numeric))
}
```

Let's run the simulation for just one parameter set:

```{r sir_example}
result <- sir(list(
  R0 = 2.0,
  infectious_period = 1.0,
  I0 = 1e-4,
  max_time = 30,
  n_points = 1000
))

result |>
  head() |>
  print()

result |>
  as_tibble() |>
  select(time, I) |>
  mutate(across(everything(), as.numeric)) |>
  ggplot(aes(time, I)) +
  geom_line()
```

If we wanted to try a simulation over multiple values, it would be easier if we
just pulled from an experiment specification. Here's the raw specification:

```{r spec}
readLines("sir_params.yaml") |>
  paste(collapse = "\n") |>
  cat()
```

Note the use of `!!float` to force YAML to parse `1e-4` as a number and not the
string `"1e-4"`.

Let's read this specification in:

```{r read_spec}
parameter_sets <- read_griddle("sir_params.yaml")
cat(yaml::as.yaml(parameter_sets))
```

The griddle produces a list of 4 parameter sets, each of which is
a list of parameter names and values.

We can run an experiment, using those parameter sets and a simulation function. Here
I call the `sir()` function again:

```{r run_experiment}
output <- run(
  fun = sir,
  parameter_sets = parameter_sets
)

str(output)
```

There is a bunch of stuff in here:

- `fun`: The simulation function
- `parameter_map`: A named list. The list elements are the parameter sets
  (which are themselves named lists of parameters). The list names are hashes
  of those parameter sets.
- `simulations`: One simulation per parameter set. Each simulation has:
  - `parameter_hash`: the hash corresponding to the parameter used
  - `result`: the output from `fun` when called on this parameter set

For convenience, there is a way to "flatten" all the results into a single
tibble, annotated with the parameter values. This works if the simulation results
are all tibbles and the parameter values are all scalars:

```{r flatten}
flatten_run(output)
```

The flat results make simple EDA easy:

```{r plot, fig.width=6}
flatten_run(output) |>
  mutate(across(c(R0, infectious_period, I0), factor)) |>
  ggplot(aes(time, I)) +
  facet_wrap(vars(infectious_period), labeller = label_both) +
  geom_line(aes(color = R0, linetype = I0)) +
  ggtitle("SIR ODE, for varying R0 and infectious period")
```

Note that, because we nested $I_0$ under $R_0$, those two values very together
(i.e., all the solid lines are blue, dashed red).
