---
title: "Stochastic simulations with griddleR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Stochastic simulations with griddleR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(griddleR)
library(ggplot2)
```

See the SIR ODE example for using a grid of values. Now we'll combine that with
replicates of stochastic simulations.

Imagine we were interested in a grid search over the shape and scale parameters
of a gamma distribution:

```{r show_spec}
readLines("stochastic_params.yaml") |>
  paste(collapse = "\n") |>
  cat()
```

For a trivial example, I'll draw just one sample from the distribution, for each
combination of parameters:

```{r experiment}
parameter_sets <- griddleR::read_griddle("stochastic_params.yaml")

f1 <- replicated(function(ps) {
  tibble::tibble(x = rgamma(1, shape = ps$shape, scale = ps$scale))
})

output <- run(fun = f1, parameter_sets = parameter_sets)

flatten_run(output)
```

We get the same results if we do that all over again:

```{r experiment_repeat}
run(fun = f1, parameter_sets = parameter_sets) |>
  flatten_run()
```

It's a bit silly to take only a single random value per seed, so instead let's
draw many:

```{r plot, fig.width = 6, fig.height = 6}
f2 <- replicated(function(ps) {
  tibble::tibble(x = rgamma(1e3, shape = ps$shape, scale = ps$scale))
})

output <- run(fun = f2, parameter_sets = parameter_sets)

flatten_run(output) |>
  ggplot(aes(factor(replicate), x)) +
  facet_wrap(vars(shape, scale), scales = "free_x", labeller = label_both) +
  geom_violin()
```

Note that the results are consistent for the same shape and scale parameters,
but they are not identical across seeds.
